---
title: "Crime Trends Over Time"
author: "Linshan - lx2346" 
output: 
  html_document:
    code_folding: hide
    always_allow_html: true
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
library(tidyverse)
library(lubridate)
library(plotly)

theme_set(theme_minimal()+theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis")

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

```{r include=FALSE}
data = read_csv("./data/crime_data_LA.csv", na = c("NA", ".", ""))
data = data |>
  select(dr_no ,date = date_occ, time = time_occ, area = area_name, severity = part_1_2, crime_type = crm_cd_desc, premis = premis_desc) |>
  mutate(
    severity = case_when(
    severity == 1 ~ "serious",
    severity == 2 ~ "less serious"
  ) )|>
    mutate(severity = factor(severity, levels = c("less serious", "serious"), ordered = TRUE),
           crime_type = tolower(crime_type))

data = data |>
  mutate(date = as.Date(sub(" .*", "", data$date), format = "%m/%d/%Y"))
data$time = sprintf("%04s", as.character(data$time))
data$time = format(strptime(data$time, format="%H%M"), "%H:%M")
```

Explore the patterns and trends in crime occurrences based on various time scales (yearly, seasonal, weekdays, and hourly). This web page allows you to visualize how crime evolves over time and across different dimensions.

## Overall Trend 
First, let's take a look at the time trend chart of the overall case distribution!
```{r include=FALSE}
data = data |>
  mutate(
    year_month = floor_date(data$date, unit = "month"),
    month = month(date),
    season = case_when(
      month %in% 3:5 ~ "Spring",
      month %in% 6:8 ~ "Summer",
      month %in% 9:11 ~ "Fall",
      month %in% c(12,1,2) ~ "Winter",
      TRUE ~ "Unknown"
    ),
    hour_period = sub(":(.*)", "", data$time),
    year = year(date),
    weekday = wday(date, label = TRUE, abbr = FALSE)
  )
```

```{r warning=FALSE, message=FALSE}
data |>
  group_by(year_month, severity) |>
  summarise(daily_count = n()) |>
  ggplot(aes(x = year_month, y = daily_count, color = severity)) +
  geom_line(size = 1) +
  labs(title = "Trends in Severe and Non-Severe Crimes",
       x = "Time", 
       y = "Crime Count") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5)) 

```

## Annual Trends
### Yearly Crime Counts
\
This section shows how the frequency of crimes has changed year by year.\
```{r warning=FALSE, message=FALSE}
data |>
  filter(year != 2024) |>
  group_by(year) |>
  summarize(count = n()) |>
  knitr::kable()
```

```{r warning=FALSE, message=FALSE, fig.width = 8, fig.asp = 0.5, out.width = "90%"}
data |>
  filter(year != 2024) |>
  group_by(year) |>
  summarize(year_count = n()) |>
  ggplot(aes(x = as.factor(year), y = year_count, fill = as.factor(year))) +
  geom_bar(stat = "identity", width = 0.5, alpha = 0.7) +  
  geom_text(aes(label = year_count), vjust = -0.5, size = 3) +
  labs(title = "Yearly Crime Counts",
       x = "Year",
       y = "Number of Crimes") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5)) 
```
\
The highest number of cases occurred in LA in **2022**.

### Serious vs. Less Serious Crimes

```{r warning=FALSE, message=FALSE, fig.width = 8, fig.asp = 0.5, out.width = "90%"}
data |>
  filter(year != 2024) |>
  group_by(year, severity) |>
  summarize(year_count = n()) |>
  ggplot(aes(x = as.factor(year), y = year_count, fill = severity)) +
  geom_bar(stat = "identity", position = "stack", width = 0.5, alpha = 0.7) +  
  geom_text(aes(label = year_count, group = severity), position = position_stack(vjust = 0.5)) +
  labs(title = "Yearly Crime Counts",
       x = "Year",
       y = "Number of Crimes") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5)) 
  
```
\
Use chi-squared test to check if there is a significant relationship between year and crime severity.
```{r warning=FALSE, message=FALSE}
year_ser = data |>
  filter(year != 2024) |>
  group_by(year, severity) |>
  summarize(year_count = n()) |>
  pivot_wider(
    names_from = year,
    values_from = year_count
  )
knitr::kable(year_ser)

data_matrix = as.table(as.matrix(year_ser[, -1]))
chi_result = chisq.test(data_matrix)
chi_result
```
\
This indicates that there are significant differences in the distribution of severity of cases between years. This means that there are differences in the proportion of cases occurring in terms of severity between years. Across all years, 2023 is more likely to have serious cases, while 2021 is more likely to have less serious cases.

### Crime by Area
Now, an interactive line graph is used to show the change in the number of cases occurring in different and years for different case types.
```{r warning=FALSE, message=FALSE}
area_year = data |>
  filter(year != 2024) |>
  group_by(year, area) |>
  summarize(year_count = n()) |>
  ggplot(aes(x = as.factor(year), y = year_count, color = area, group = area)) +
  geom_line(size = 0.5, alpha = 0.5) +  
  geom_point(size = 0.7) +
  labs(title = "Crime Counts by Area Over Time",
       x = "Year",
       y = "Number of Crimes",
       color = "Area") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))

ggplotly(area_year)
  
```
From the general trend observed in the line chart, the lines show an overall upward trend from 2021 to 2022, followed by a downward trend from 2022 to 2023. This indicates that the number of crimes in 2022 increased compared to 2020, 2021, and 2023. Particularly in the Central area, the number of crimes in 2022 saw a significant increase compared to 2021. Regarding the frequency of incidents, there are also significant differences between areas. Areas like `Central`, `77 Street`, and `Pacific` report a high number of crimes each year, with annual crime counts exceeding 10,000, and the peak years even reaching up to 17,000 incidents. In contrast, areas like `Devonshire`, `Hollenbeck`, and `Mission` experience fewer crimes each year, with counts typically ranging from 7,500 to 9,000 incidents.

### Crime Types Over Year
```{r warning=FALSE, message=FALSE}
crime_year = data |>
  filter(year != 2024) |>
  group_by(crime_type, year) |>
  summarize(year_count = n()) |>
  mutate(year_count = ifelse(is.na(year_count), 0, year_count)) |>
  ggplot(aes(x = year, y = year_count, color = crime_type, group = crime_type)) +
  geom_line(size = 0.3, alpha = 0.5) +  
  geom_point(size = 0.5) +
  labs(title = "Crime Counts by Area Over Time",
       x = "Year",
       y = "Number of Crimes",
       color = "Area") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5)) +
  theme(legend.position = "none") 

ggplotly(crime_year)
```
\
As shown in the figure, the most frequent crime categories are `vehicle stolen`, `battery - simple assault`, and `identity theft`. The time trends for different crime types vary across years. While most crime types have a relatively similar number of incidents each year, there are also types that show a steady increase in incidents annually (such as `vehicle stolen` and `petty theft`). Additionally, there are types that saw a sudden spike in 2022, but quickly decreased (such as `battery - simple assault`).

## Seasonal Trends
### Crime Counts By Season
```{r warning=FALSE, message=FALSE}
data |>
  filter(year != 2024) |>
  group_by(season) |>
  summarize(count = n()) |>
  knitr::kable()
```

```{r warning=FALSE, message=FALSE, fig.width = 8, fig.asp = 0.5, out.width = "90%"}
data |>
  filter(year != 2024) |>
  group_by(season) |>
  summarize(season_count = n()) |>
  ggplot(aes(x = as.factor(season), y = season_count, fill = season)) +
  geom_bar(stat = "identity", width = 0.5, alpha = 0.7) +  
  geom_text(aes(label = season_count), vjust = -0.5, size = 3) +
  labs(title = "Crime Counts By Season",
       x = "Season",
       y = "Number of Crimes") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))
```

### Serious vs. Less Serious Crimes

```{r warning=FALSE, message=FALSE, fig.width = 8, fig.asp = 0.5, out.width = "90%"}
data |>
  filter(year != 2024) |>
  mutate(season = factor(season, levels = c("Spring", "Summer", "Fall", "Winter"))) |>
  group_by(season, severity) |>
  summarize(season_count = n()) |>
  ggplot(aes(x = as.factor(season), y = season_count, fill = severity)) +
  geom_bar(stat = "identity", position = "stack", width = 0.5, alpha = 0.7) +  
  geom_text(aes(label = season_count, group = severity), position = position_stack(vjust = 0.5)) +
  labs(title = "Crime Counts By Season and Severity",
       x = "Season",
       y = "Number of Crimes") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))
  
```
\
Use chi-squared test to check if there is a significant relationship between year and crime severity.
```{r warning=FALSE, message=FALSE}
season_ser = data |>
  filter(year != 2024) |>
  mutate(season = factor(season, levels = c("Spring", "Summer", "Fall", "Winter"))) |>
  group_by(season, severity) |>
  summarize(season_count = n()) |>
  pivot_wider(
    names_from = season,
    values_from = season_count
  )
knitr::kable(season_ser)

data_matrix = as.table(as.matrix(season_ser[, -1]))
chi_result = chisq.test(data_matrix)
chi_result
```
\
This indicates that there are significant differences in the distribution of severity of cases between seasons. This means that there are differences in the proportion of cases occurring in terms of severity between seasons. Across four seasons, **Fall** is more likely to have serious cases, while **Spring** is more likely to have less serious cases.

### Crime by Area
Now, a line graph is used to show the change in the number of cases occurring in different seasons and different case types.
```{r warning=FALSE, message=FALSE}
area_season = data |>
  filter(year != 2024) |>
  mutate(season = factor(season, levels = c("Spring", "Summer", "Fall", "Winter"))) |>
  group_by(season, area) |>
  summarize(season_count = n()) |>
  ggplot(aes(x = as.factor(season), y = season_count, color = area, group = area)) +
  geom_line(size = 0.5, alpha = 0.5) +  
  geom_point(size = 0.7) +
  labs(title = "Crime Counts by Area Over Season",
       x = "Season",
       y = "Number of Crimes",
       color = "Area") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))

ggplotly(area_season)
  
```
Among the four seasons, most areas experience more cases in **summer** or **autumn**, which may be due to the warmer weather in these seasons, leading to more outdoor activities. In other regions, the number of cases is relatively balanced throughout the four seasons. The areas with the highest number of cases are `Central`, `77th Street`, `Pacific`, and `Southwest`.

### Crime Types Over Seasons
```{r warning=FALSE, message=FALSE}
crime_year = data |>
  filter(year != 2024) |>
  mutate(season = factor(season, levels = c("Spring", "Summer", "Fall", "Winter"))) |>
  group_by(crime_type, season) |>
  summarize(season_count = n()) |>
  mutate(season_count = ifelse(is.na(season_count), 0, season_count)) |>
  ggplot(aes(x = season, y = season_count, color = crime_type, group = crime_type)) +
  geom_line(size = 0.3, alpha = 0.5) +  
  geom_point(size = 0.5) +
  labs(title = "Crime Counts by Crime Type By Season",
       x = "Season",
       y = "Number of Crimes",
       color = "Area") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5)) +
  theme(legend.position = "none")

ggplotly(crime_year)
```
Overall, most crime types have relatively stable occurrences throughout the four seasons. Notable seasonal trends include `vehicle stolen`, which peaks in **autumn**, `battery - simple assault` and `assault with a deadly weapon`, which are most frequent in **summer**. Among all crime types, `vehicle stolen` and `battery - simple assault` stand out with significantly higher case numbers compared to other types.

## Weekday Trends
### Crime Counts By Weekday
```{r warning=FALSE, message=FALSE}
data |>
  group_by(weekday) |>
  summarize(count = n()) |>
  knitr::kable()
```


```{r warning=FALSE, message=FALSE, fig.width = 8, fig.asp = 0.5, out.width = "90%"}
data |>
  group_by(weekday) |>
  summarize(weekday_count = n()) |>
  ggplot(aes(x = as.factor(weekday), y = weekday_count, fill = weekday)) +
  geom_bar(stat = "identity", width = 0.5, alpha = 0.7) +  
  geom_text(aes(label = weekday_count), vjust = -0.5, size = 3) +
  labs(title = "Crime Counts By Weekday",
       x = "Weekdays",
       y = "Number of Crimes") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))
```
\
The number of crimes on **Fridays** and **Saturdays** is significantly higher than on other days of the week. This is not hard to understand, as people usually attend parties, nightlife, or other social activities on Fridays and Saturdays, which may increase the likelihood of violent incidents (such as fights or assaults) or thefts.

### Serious vs. Less Serious Crimes

```{r warning=FALSE, message=FALSE, fig.width = 8, fig.asp = 0.5, out.width = "90%"}
data |>
  group_by(weekday, severity) |>
  summarize(weekday_count = n()) |>
  ggplot(aes(x = as.factor(weekday), y = weekday_count, fill = severity)) +
  geom_bar(stat = "identity", position = "stack", width = 0.5, alpha = 0.7) +  
  geom_text(aes(label = weekday_count, group = severity), position = position_stack(vjust = 0.5)) +
  labs(title = "Crime Counts By Weekday and Severity",
       x = "Weekdays",
       y = "Number of Crimes") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))
  
```
\
Use chi-squared test to check if there is a significant relationship between weekdays and crime severity.
```{r warning=FALSE, message=FALSE}
weekday_ser = data |>
  group_by(weekday, severity) |>
  summarize(weekday_count = n()) |>
  pivot_wider(
    names_from = weekday,
    values_from = weekday_count
  )
knitr::kable(weekday_ser)

data_matrix = as.table(as.matrix(weekday_ser[, -1]))
chi_result = chisq.test(data_matrix)
chi_result
```
\
This indicates that there are significant differences in the distribution of severity of cases between weekdays. This means that there are differences in the proportion of cases occurring in terms of severity between weekdays. Across these weekdays, **Saturday and Friday** have more crime cases and are more likely to have serious cases.

### Crime by Area
Now, a line graph is used to show the change in the number of cases occurring in different weekdays and different case types.
```{r warning=FALSE, message=FALSE}
area_weekday = data |>
  group_by(weekday, area) |>
  summarize(weekday_count = n()) |>
  ggplot(aes(x = as.factor(weekday), y = weekday_count, color = area, group = area)) +
  geom_line(size = 0.5, alpha = 0.5) +  
  geom_point(size = 0.7) +
  labs(title = "Crime Counts by Area Over Weekday",
       x = "Weekdays",
       y = "Number of Crimes",
       color = "Area") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))

ggplotly(area_weekday)
  
```
\
The areas with the highest number of cases are `Central`, `77th Street`, `Pacific`, and `Southwest`. Overall, from the perspective of weekday trends, the number of crimes decreases from Sunday to Tuesday, remains relatively stable on Wednesday and Thursday, and then shows a significant increase on Friday and Saturday.

### Crime Types Over Weekdays
```{r warning=FALSE, message=FALSE}
crime_weekday = data |>
  group_by(crime_type, weekday) |>
  summarize(weekday_count = n()) |>
  mutate(weekday_count = ifelse(is.na(weekday_count), 0, weekday_count)) |>
  ggplot(aes(x = weekday, y = weekday_count, color = crime_type, group = crime_type)) +
  geom_line(size = 0.3, alpha = 0.5) +  
  geom_point(size = 0.5) +
  labs(title = "Crime Counts by Crime Type By Weekday",
       x = "Weekdays",
       y = "Number of Crimes") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5)) +
  theme(legend.position = "none")

ggplotly(crime_weekday)
```
\
Overall, four types of time trends were observed:\
1. The highest number of cases occurred on Fridays and Saturdays, with a sharp increase on Friday, a slight decrease on Saturday, and a return to baseline on Sunday. Examples include `vehicle stolen`, `burglary, and petty theft (under $950)`.\
2. The highest number of cases occurred on Saturdays and Sundays, while the rest of the week remained near the baseline. Examples include `battery - simple assault` and `intimate partner - simple assault`.\
3. The lowest number of cases occurred on Saturdays and Sundays, with the rest of the week remaining near the baseline. An example is `identity theft`.\
4. The number of cases was similar across all days of the week, with minimal fluctuation.

## Hourly Trends
### Crime Counts By Hour of The Day

```{r warning=FALSE, message=FALSE, fig.width = 8, fig.asp = 0.5, out.width = "90%"}
count_hour = data |>
  mutate(hour_period = as.factor(hour_period)) |>
  group_by(hour_period) |>
  summarize(hour_count = n()) |>
  ggplot(aes(x = hour_period, y = hour_count, fill = hour_period)) +
  geom_bar(stat = "identity", width = 0.5, alpha = 0.7) +  
  labs(title = "Crime Counts By The Hour Of The Day",
       x = "Hour Periods",
       y = "Number of Crimes") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))

ggplotly(count_hour)
```
\
The highest number of cases occurs **at noon(12:00)** and in the **evening(18:00/19:00)**.

### Serious vs. Less Serious Crimes

```{r warning=FALSE, message=FALSE, fig.width = 8, fig.asp = 0.5, out.width = "90%"}
ser_hour = data |>
  mutate(hour_period = as.factor(hour_period)) |>
  group_by(hour_period, severity) |>
  summarize(hour_count = n()) |>
  ggplot(aes(x = hour_period, y = hour_count, fill = severity)) +
  geom_bar(stat = "identity", position = "stack", width = 0.5, alpha = 0.7) +  
  labs(title = "Crime Counts By Hour Of The Day",
       x = "Hour Periods",
       y = "Number of Crimes") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))

ggplotly(ser_hour)
```
\
\
Use chi-squared test to check if there is a significant relationship between hour period of the day and crime severity.
```{r warning=FALSE, message=FALSE}
hour_ser = data |>
  group_by(hour_period, severity) |>
  summarize(hour_count = n()) |>
  pivot_wider(
    names_from = hour_period,
    values_from = hour_count
  )

first_half = hour_ser[,1:13]
second_half = hour_ser[,c(1, 14:25) ]

knitr::kable(first_half) 
knitr::kable(second_half)

data_matrix = as.table(as.matrix(hour_ser[, -1]))
chi_result = chisq.test(data_matrix)
chi_result
```
\
This indicates that there are significant differences in the distribution of severity of cases in different hour periods of the day, which means that there are differences in the proportion of cases occurring in terms of severity. **Across 24 hours of a day, Around noon, the number of cases increases, but less serious incidents are more likely to occur. As the evening progresses into midnight, although the total number of cases decreases, serious incidents become more likely.**

### Crime by Area
Now, a line graph is used to show the change in the number of cases occurring in different hour periods of the day and different case types.
```{r warning=FALSE, message=FALSE}
area_hour = data |>
  group_by(hour_period, area) |>
  summarize(hour_count = n()) |>
  ggplot(aes(x = as.factor(hour_period), y = hour_count, color = area, group = area)) +
  geom_line(size = 0.4, alpha = 0.5) +  
  geom_point(size = 0.6) +
  labs(title = "Crime Counts by Area Over Hour Period of The Day",
       x = "Hour Periods",
       y = "Number of Crimes",
       color = "Area") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))

ggplotly(area_hour)
  
```
\
It can be observed that the distribution of cases by region is similar to the overall distribution of cases. **Two peaks** are evident: the first is a **sharp peak occurring at 12:00**, while the second peak, with its highest point around **18:00–19:00** in the evening, forms a gradual slope spanning across a broader time range.

### Crime Types By Hour Period of The Day
```{r warning=FALSE, message=FALSE}
crime_hour = data |>
  group_by(crime_type, hour_period) |>
  summarize(hour_count = n()) |>
  mutate(hour_count = ifelse(is.na(hour_count), 0, hour_count)) |>
  ggplot(aes(x = hour_period, y = hour_count, color = crime_type, group = crime_type)) +
  geom_line(size = 0.3, alpha = 0.5) +  
  geom_point(size = 0.5) +
  labs(title = "Crime Counts by Crime Type By Hour Period of The Day",
       x = "Hour Periods",
       y = "Number of Crimes") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5)) +
  theme(legend.position = "none")

ggplotly(crime_hour)
```
\
When categorized by crime type, it is evident that the peak times vary across different types of crimes. `Identity theft` shows a pronounced peak at noon (12:00), while `vehicle theft` or `burglary from vehicles` is more likely to occur during the evening hours (16:00–23:00).




